% $Id: ldap-controls.tex,v 1.3 2008/03/20 12:24:57 stroeder Exp $

\section{\module{ldap.controls} ---
  High-level access to LDAP controls }

\declaremodule{standard}{ldap.controls}

% Author of the module code;
\moduleauthor{python-ldap project}{python-ldap-dev@lists.sourceforge.net}

\modulesynopsis{High-level access to LDAP controls.}

The \module{ldap.controls} module defines the following classes:

\begin{classdesc}{LDAPControl}{controlType, criticality \optional{, controlValue=\constant{None} \optional{, encodedControlValue=\constant{None}}}}
Base class for all LDAP controls. This class should not be used directly,
instead one of the following subclasses should be used as appropriate.

\begin{methoddesc}[LDAPControl]{encodeControlValue}{value}
Dummy method to be overridden by subclasses.
\end{methoddesc}

\begin{methoddesc}[LDAPControl]{decodeControlValue}{value}
Dummy method to be overridden by subclasses.
\end{methoddesc}

\begin{methoddesc}[LDAPControl]{getEncodedTuple}{}
Return a readily encoded 3-tuple which can be directly
passed to C module _ldap. This method is called by
function \function{ldap.EncodeControlTuples}.
\end{methoddesc}
\end{classdesc}

\begin{classdesc}{BooleanControl}{controlType, criticality \optional{, controlValue=\constant{None} \optional{, encodedControlValue=\constant{None}}}}
Base class for simple controls with booelan control value.

In this base class \var{controlValue} has to be passed as
boolean type (\constant{True}/\constant{False} or \constant{1}/\constant{0}).
\end{classdesc}
\begin{classdesc}{SimplePagedResultsControl}{controlType, criticality \optional{, controlValue=\constant{None} \optional{, encodedControlValue=\constant{None}}}}
The class provides the LDAP Control Extension for Simple Paged Results Manipulation. \var{controlType} is ignored
in favor of \constant{ldap.LDAP_CONTROL_PAGE_OID}.
\begin{seealso}
\seerfc{2696}{LDAP Control Extension for Simple Paged Results Manipulation}{}
\end{seealso}
\end{classdesc}

\begin{classdesc}{MatchedValuesControl}{criticality \optional{, controlValue=\constant{None}}}
This class provides the LDAP Matched Values control. \var{controlValue} is an LDAP filter.
\begin{seealso}
\seerfc{3876}{Returning Matched Values with the Lightweight Directory Access Protocol version 3 (LDAPv3)}{}
\end{seealso}
\end{classdesc}

The \module{ldap.controls} module defines the following functions:

\begin{funcdesc}{EncodeControlTuples}{ldapControls} % -> list
  Returns list of readily encoded 3-tuples which can be directly
  passed to C module _ldap.
\end{funcdesc}


\begin{funcdesc}{DecodeControlTuples}{ldapControlTuples} % -> list
  Decodes a list of readily encoded 3-tuples as returned by the C module _ldap.
\end{funcdesc}

