% $Id: ldap-modlist.tex,v 1.3 2008/03/20 12:24:57 stroeder Exp $

\section{\module{ldap.modlist} ---
  Generate modify lists }

\declaremodule{standard}{ldap.modlist}

% Author of the module code;
\moduleauthor{python-ldap project}{python-ldap-dev@lists.sourceforge.net}

\modulesynopsis{Generate modify lists.}

The \module{ldap.modlist} module defines the following functions:

\begin{funcdesc}{addModlist}{entry \optional{, ignore_attr_types=\constant{[]}}} % -> list

This function builds a list suitable for passing it
directly as argument \var{modlist} to method \method{add()} or
its synchronous counterpart \method{add_s()}.

\var{entry} is a dictionary like returned when
receiving search results.

\end{funcdesc}

\begin{funcdesc}{modifyModlist}{ old_entry, new_entry \optional{, ignore_attr_types=\constant{[]} \optional{, ignore_oldexistent=\constant{0}}}} % -> list

This function builds a list suitable for passing it directly as argument
\var{modlist} to method \method{modify()} or its synchronous counterpart
\method{modify_s()}.

Roughly when applying the resulting modify list to an entry holding
the data \var{old_entry} it will be modified in such a way that the
entry holds \var{new_entry} after the modify operation. It is handy in
situations when it is impossible to track user changes to an entry's
data or for synchronizing operations.

\var{old_entry} and \var{new_entry} are dictionaries
like returned when receiving search results.

\var{ignore_attr_types} is a list of attribute type names which
shall be ignored completely. These attribute types will not appear
in the result.

If \var{ignore_oldexistent} is non-zero attribute type names which
are in \var{old_entry} but are not found in \var{new_entry} at all
are not deleted. This is handy for situations where your application
sets attribute value to '' for deleting an attribute.
In most cases leave zero.

\end{funcdesc}

