
% $Id: constants.tex,v 1.3 2000/08/20 14:14:27 leonard Exp $

\section{Exceptions}

The following exceptions and constants are exported from the module:

%\renewcommand{\indexsubitem}{(in module ldap)}
\newcommand{\excline}[1]{\item[\bfcode{#1}\ttindex{#1}]}

%%------------------------------------------------------------
%% error

\begin{excdesc}{LDAPError}
This is the base class of all execeptions raised when an error arises within
the glue code between the \C\ interface and the Python API.
When raised, the associated value is either a dictionary of descriptions,
or a simple string.
\end{excdesc}

%%------------------------------------------------------------
%% ALIAS_DEREF_PROBLEM ... USER_CANCELLED

\begin{excdesc}{ALIAS_DEREF_PROBLEM}
	A  problem was encountered when dereferencing an alias.
\end{excdesc}
\begin{excdesc}{ALIAS_PROBLEM}
	An alias in the directory points to a nonexistent entry.
\end{excdesc}
\begin{excdesc}{ALREADY_EXISTS}
	The entry already exists.
\end{excdesc}
\begin{excdesc}{AUTH_UNKNOWN}
	The authentication method specified to \method{bind} is not known.
\end{excdesc}
\begin{excdesc}{BUSY}
	The DSA is busy.
\end{excdesc}
\begin{excdesc}{COMPARE_FALSE}
	A compare operation returned false.
	(This exception should never be seen.)
\end{excdesc}
\begin{excdesc}{COMPARE_TRUE}
	A compare operation returned true.
	(This exception should never be seen.)
\end{excdesc}
\begin{excdesc}{CONSTRAINT_VIOLATION}
	An attribute value specified  violates
	some constraint (e.g., a postalAddress
	has too many lines, or a line that  is
	too long).
\end{excdesc}
\begin{excdesc}{DECODING_ERROR}
	An  error  was  encountered decoding a
	result from the LDAP server.
\end{excdesc}
\begin{excdesc}{ENCODING_ERROR}
	An   error  was  encountered  encoding
	parameters to send to the LDAP server.
\end{excdesc}
\begin{excdesc}{FILTER_ERROR}
	An  invalid  filter  was  supplied  to
	\method{search} (e.g., unbalanced parentheses).
\end{excdesc}
\begin{excdesc}{INAPPROPRIATE_AUTH}
	Inappropriate authentication was spec-
	ified   (e.g.,   \constant{LDAP_AUTH_SIMPLE} was
	specified and the entry does not  have
	a userPassword attribute).
\end{excdesc}
\begin{excdesc}{INAPPROPRIATE_MATCHING}
	   Filter  type  not  supported  for  the specified attribute.
\end{excdesc}
\begin{excdesc}{INSUFFICIENT_ACCESS}
	   The user has  insufficient  access  to perform the operation.
\end{excdesc}
\begin{excdesc}{INVALID_CREDENTIALS}
	Invalid   credentials  were  presented during \method{bind}.
	(e.g., the wrong password).
\end{excdesc}
\begin{excdesc}{INVALID_DN_SYNTAX}
	   A syntactically invalid DN was  specified.
\end{excdesc}
\begin{excdesc}{INVALID_SYNTAX}
	An  invalid attribute value was specified.
\end{excdesc}
\begin{excdesc}{IS_LEAF}
	The object specified is a leaf of the diretcory tree.
\end{excdesc}
\begin{excdesc}{LOCAL_ERROR}
	Some local error  occurred.
	This is usually due to failed memory allocation.
\end{excdesc}
\begin{excdesc}{LOOP_DETECT}
	A loop was detected.
\end{excdesc}
\begin{excdesc}{NAMING_VIOLATION}
	A naming violation occurred.
\end{excdesc}
\begin{excdesc}{NOT_ALLOWED_ON_NONLEAF}
	The operation is not allowed on a non-leaf object.
\end{excdesc}
\begin{excdesc}{NOT_ALLOWED_ON_RDN}
	The  operation  is  not  allowed on an RDN.
\end{excdesc}
\begin{excdesc}{NO_OBJECT_CLASS_MODS}
	Object  class  modifications  are  not allowed.
\end{excdesc}
\begin{excdesc}{NO_SUCH_ATTRIBUTE}
	The  attribute type specified does not exist in the entry.
\end{excdesc}
\begin{excdesc}{NO_SUCH_OBJECT}
	The specified object does not exist in the directory.
\end{excdesc}
\begin{excdesc}{OBJECT_CLASS_VIOLATION}
	An  object  class  violation  occurred (e.g.,  
	a "must" attribute was missing from the entry).
\end{excdesc}
\begin{excdesc}{OPERATIONS_ERROR}
	An operations error occurred.
\end{excdesc}
\begin{excdesc}{OTHER}
	An unclassified error occurred.
\end{excdesc}
\begin{excdesc}{PARAM_ERROR}
	An ldap routine was called with a  bad parameter.
\end{excdesc}
\begin{excdesc}{PARTIAL_RESULTS}
	Partial results only returned. 
	(This exception should never be seen.)
\end{excdesc}
\begin{excdesc}{PROTOCOL_ERROR}
	A violation of the LDAP protocol was detected.
\end{excdesc}
\begin{excdesc}{RESULTS_TOO_LARGE}
	(Undocumented.)
\end{excdesc}
\begin{excdesc}{SERVER_DOWN}
	The  LDAP  library  can't  contact the LDAP server.
\end{excdesc}
\begin{excdesc}{SIZELIMIT_EXCEEDED}
	An LDAP size limit was exceeded.
\end{excdesc}
\begin{excdesc}{STRONG_AUTH_NOT_SUPPORTED}
	The   LDAP  server  does  not  support strong authentication.
\end{excdesc}
\begin{excdesc}{STRONG_AUTH_REQUIRED}
	Strong authentication is required  for the operation.
\end{excdesc}
\begin{excdesc}{TIMELIMIT_EXCEEDED}
	An LDAP time limit was exceeded.
\end{excdesc}
\begin{excdesc}{TIMEOUT}
	A timelimit was exceeded while waiting for a result from the server.
\end{excdesc}
\begin{excdesc}{TYPE_OR_VALUE_EXISTS}
	An  attribute  type or attribute value specified already 
	exists in the entry.
\end{excdesc}
\begin{excdesc}{UNAVAILABLE}
	The DSA is unavailable.
\end{excdesc}
\begin{excdesc}{UNDEFINED_TYPE}
	(Undocumented.)
\end{excdesc}
\begin{excdesc}{UNWILLING_TO_PERFORM}
	The  DSA  is  unwilling to perform the operation.
\end{excdesc}
\begin{excdesc}{USER_CANCELLED}
	The operation was cancelled via the \method{abandon} method.
\end{excdesc}
These exceptions are raised when a result code from an underlying API
call does not indicate success.
\end{excdesc}

\section{Constants}

%%------------------------------------------------------------
%% PORT

\begin{datadesc}{PORT}
The standard TCP port that LDAP servers listen on (389).
\end{datadesc}

%%============================================================
%% others

Other, undocumented constants are available in the module and
correspond to constants defined in the underlying library's
C header files.
They should not be relied upon as being available in future versions.
